---
output:
  html_document: default
  pdf_document: default
---

#1.1: Load the dataset and display the first 10 rows w/ summary

```{r}

library(conflicted)
library(tidyverse)
library(tidytext)
library(lubridate)
library(ggplot2)
library(wordcloud)
library(RColorBrewer)
library(scales)
library(igraph)
library(ggraph)
library(tm)
library(e1071)
library(caret)
library(reshape2)

conflicts_prefer(dplyr::filter)

df <- read.csv("twitter_sentiment_data.csv", stringsAsFactors = FALSE)

cat("Column names:\n")
print(names(df))

cat("\nStructure (first 10 rows):\n")
print(head(df, 10))

cat("\nSummary:\n")
print(summary(df))

```

#1.2: Report the detected columns in the dataset

```{r}

guess_text_col <- function(cols) {
  candidates <- c("text", "tweet", "content", "body", "message")
  found <- intersect(candidates, cols)
  if (length(found) > 0) return(found[1])
  return(NULL)
}

guess_sentiment_col <- function(cols) {
  candidates <- c("sentiment", "label", "sentiment_label", "polarity")
  found <- intersect(candidates, cols)
  if (length(found) > 0) return(found[1])
  return(NULL)
}

text_col      <- guess_text_col(names(df))
sentiment_col <- guess_sentiment_col(names(df))

cat("Detected text column:", text_col, "\n")
cat("Detected sentiment column:", sentiment_col, "\n")

```

#2.1 Cleaning and basic setups using the detected columns

```{r}

if (is.null(text_col)) {
  stop("No text column detected. Please set text_col manually.")
}

clean_tweets <- function(data, text_col) {
  data %>%
    mutate(
      text_clean = .data[[text_col]] %>%
        tolower() %>%
        gsub("http[s]?://\\S+", " ", .) %>%
        gsub("@\\w+", " ", .) %>%
        gsub("#", " ", .) %>%
        gsub("[^a-z0-9\\s]", " ", .) %>%
        gsub("\\s+", " ", .) %>%
        trimws()
    )
}

df <- clean_tweets(df, text_col)

data("stop_words")

make_wordcloud <- function(tokens_df, n_max = 150, min_freq = 3, title = "") {
  freq <- tokens_df %>%
    count(word, sort = TRUE) %>%
    filter(n >= min_freq) %>%
    slice_max(order_by = n, n = n_max)

  if (nrow(freq) == 0) {
    message("No words to display for: ", title)
    return(invisible(NULL))
  }

  set.seed(123)
  wordcloud(
    words       = freq$word,
    freq        = freq$n,
    scale       = c(4, 0.7),
    max.words   = n_max,
    random.order= FALSE,
    rot.per     = 0.2,
    colors      = brewer.pal(8, "Dark2")
  )
  title(main = title)
}

```

#2.2 Justify word cloud generation by constructing a helper function

```{r}

make_wordcloud <- function(tokens_df,
                           n_max   = 200,      
                           min_freq = 1,       
                           title   = "") {

  if (nrow(tokens_df) == 0) {
    message("No tokens available for: ", title)
    return(invisible(NULL))
  }

  freq <- tokens_df %>%
    dplyr::filter(!is.na(word), word != "") %>%
    count(word, sort = TRUE) %>%
    dplyr::filter(n >= min_freq, is.finite(n)) %>%
    slice_max(order_by = n, n = n_max)

  if (nrow(freq) == 0) {
    message("No words left after filtering for: ", title)
    return(invisible(NULL))
  }

  set.seed(123)
  suppressWarnings(  
    wordcloud(
      words        = freq$word,
      freq         = freq$n,
      scale        = c(3.5, 0.4),  
      max.words    = n_max,
      random.order = FALSE,
      rot.per      = 0.2,
      colors       = brewer.pal(8, "Dark2")
    )
  )
  title(main = title)
}

```

#3.1 Generating word cloud

```{r}

df_topics <- df %>%
  mutate(
    text_clean = ifelse(is.na(text_clean), "", text_clean),
    topic_media_doc = str_detect(text_clean,
      "before the flood|leonardo dicaprio|dicaprio|film|documentary"),
    topic_politics  = str_detect(text_clean,
      "trump|clinton|president|republican|democrat|policy|vote|election|government"),
    topic_science   = str_detect(text_clean,
      "data|research|scientist|evidence|temperature|co2|emission|ipcc|study"),
    topic_solutions = str_detect(text_clean,
      "renewable|solar|wind|innovation|technology|solutions?|mitigation|adaptation")
  )

tokens_media <- df_topics %>%
  dplyr::filter(topic_media_doc) %>%
  unnest_tokens(word, text_clean) %>%
  dplyr::filter(!is.na(word), word != "") %>%
  anti_join(stop_words, by = "word")

tokens_non_media <- df_topics %>%
  dplyr::filter(!topic_media_doc) %>%
  unnest_tokens(word, text_clean) %>%
  dplyr::filter(!is.na(word), word != "") %>%
  anti_join(stop_words, by = "word")

par(mfrow = c(1, 2))
make_wordcloud(tokens_media,     title = "Media/Documentary-related tweets")
make_wordcloud(tokens_non_media, title = "Non-media tweets")
par(mfrow = c(1, 1))

tokens_politics <- df_topics %>%
  dplyr::filter(topic_politics) %>%
  unnest_tokens(word, text_clean) %>%
  dplyr::filter(!is.na(word), word != "") %>%
  anti_join(stop_words, by = "word")

tokens_science <- df_topics %>%
  dplyr::filter(topic_science) %>%
  unnest_tokens(word, text_clean) %>%
  dplyr::filter(!is.na(word), word != "") %>%
  anti_join(stop_words, by = "word")

par(mfrow = c(1, 2))
make_wordcloud(tokens_politics, title = "Political discourse")

make_wordcloud(tokens_science,  title = "Scientific discourse")
par(mfrow = c(1, 1))

```

#3.2 Comparing the message (cross-sentiment analysis)

```{r}

if (!is.null(sentiment_col)) {
  sent_vals <- df[[sentiment_col]]
  if (is.numeric(sent_vals)) {
    df$sentiment_num <- sent_vals
  } else {
    df$sentiment_num <- as.numeric(as.factor(sent_vals)) - 1
  }

  df <- df %>%
    mutate(
      sentiment_cat = case_when(
        sentiment_num %in% c(0)         ~ "Negative",
        sentiment_num %in% c(1)         ~ "Neutral",
        sentiment_num %in% c(2, 3, 4)   ~ "Positive",
        TRUE                            ~ "Other"
      )
    )

  df_cross <- df %>%
    mutate(
      has_problem = str_detect(text_clean,
        "melting|flood|drought|disaster|extinction|crisis|catastrophe"),
      has_solution = str_detect(text_clean,
        "renewable|solar|wind|policy|tax|regulation|solution|innovation")
    )

  tokens_pos_solution <- df_cross %>%
    filter(sentiment_cat == "Positive", has_solution) %>%
    unnest_tokens(word, text_clean) %>%
    anti_join(stop_words, by = "word")

  tokens_neg_problem <- df_cross %>%
    filter(sentiment_cat == "Negative", has_problem) %>%
    unnest_tokens(word, text_clean) %>%
    anti_join(stop_words, by = "word")

  par(mfrow = c(1, 2))
  make_wordcloud(tokens_pos_solution, title = "Positive + Solutions")
  make_wordcloud(tokens_neg_problem,  title = "Negative + Problems")
  par(mfrow = c(1, 1))
} else {
  message("No sentiment column detected; skipping sentiment-based splits.")
}

```

#4.1 Preparation in conducting prediction (dataset cleaning)

```{r}

df_ml <- df %>%
  select(text = all_of(text_col),
         sentiment = all_of(sentiment_col)) %>%
  filter(!is.na(text), !is.na(sentiment))

df_ml$sentiment <- as.factor(df_ml$sentiment)

clean_fun <- function(x) {
  x <- tolower(x)
  x <- gsub("http[s]?://\\S+", " ", x)   
  x <- gsub("@\\w+", " ", x)            
  x <- gsub("#", " ", x)                
  x <- gsub("[^a-z0-9\\s]", " ", x)     
  x <- gsub("\\s+", " ", x)            
  trimws(x)
}

df_ml$text_clean <- clean_fun(df_ml$text)

set.seed(123)

train_index <- createDataPartition(df_ml$sentiment, p = 0.8, list = FALSE)

train_df <- df_ml[train_index, ]
test_df  <- df_ml[-train_index, ]

```

#4.2 Creating a document–term matrix (DTM)

```{r}

train_corpus <- VCorpus(VectorSource(train_df$text_clean))
test_corpus  <- VCorpus(VectorSource(test_df$text_clean))

to_space <- content_transformer(function(x, pattern) gsub(pattern, " ", x))

train_corpus <- train_corpus %>%
  tm_map(removeNumbers) %>%
  tm_map(removeWords, stopwords("en")) %>%
  tm_map(stripWhitespace)

test_corpus <- test_corpus %>%
  tm_map(removeNumbers) %>%
  tm_map(removeWords, stopwords("en")) %>%
  tm_map(stripWhitespace)

dtm_train <- DocumentTermMatrix(train_corpus)
dtm_train <- removeSparseTerms(dtm_train, 0.98)  
dtm_train

dtm_test <- DocumentTermMatrix(test_corpus,
                               list(dictionary = Terms(dtm_train)))

train_matrix <- as.data.frame(as.matrix(dtm_train))
test_matrix  <- as.data.frame(as.matrix(dtm_test))

train_matrix <- train_matrix %>% setNames(make.names(names(.)))
test_matrix  <- test_matrix  %>% setNames(make.names(names(.)))

train_matrix$sentiment <- train_df$sentiment
test_labels            <- test_df$sentiment

dim(train_matrix)

```

#4.3 Run and train the Naive Bayes (NB) classifier

```{r}

set.seed(123)

nb_model <- naiveBayes(
  sentiment ~ .,
  data = train_matrix,
  laplace = 1
)
nb_model
```

#4.4 Evaluation on test set (in numerical value)

```{r}

pred_test <- predict(nb_model, newdata = test_matrix)

conf_mat <- confusionMatrix(pred_test, test_labels)
conf_mat

```

#4.5 Confusion Matrix heatmap

```{r}

conf_mat <- confusionMatrix(pred_test, test_labels)

cm_df <- as.data.frame(conf_mat$table)
colnames(cm_df) <- c("Reference", "Prediction", "Freq")
ggplot(cm_df, aes(x = Reference, y = Prediction, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Freq), color = "black", size = 4) +
  scale_fill_gradient(low = "#f7fbff", high = "#08306b") +
  coord_equal() +
  labs(
    title = "Confusion Matrix (Test Set)",
    x = "True Sentiment",
    y = "Predicted Sentiment",
    fill = "Count"
  ) +
  theme_minimal()

```

#4.6 Overall correctness (accuracy vs error) barplot

```{r}

overall_acc <- conf_mat$overall["Accuracy"]
overall_err <- 1 - overall_acc

acc_df <- data.frame(
  Metric = c("Correct", "Incorrect"),
  Value  = c(overall_acc, overall_err)
)

ggplot(acc_df, aes(x = Metric, y = Value, fill = Metric)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 1)) +
  scale_fill_manual(values = c("Correct" = "#2ca25f", "Incorrect" = "#de2d26")) +
  labs(
    title = "Overall Classification Accuracy",
    x = NULL,
    y = "Proportion"
  ) +
  theme_minimal()

```

#4.7 Reporting the Per-class precision / recall / F1 of the test set (in bar chart)

```{r}

cm <- conf_mat$table          
classes <- colnames(cm)       

metrics_list <- lapply(seq_along(classes), function(i) {
  cls <- classes[i]
  tp <- cm[cls, cls]
  fp <- sum(cm[, cls]) - tp
  fn <- sum(cm[cls, ]) - tp
  tn <- sum(cm) - tp - fp - fn
  precision <- if ((tp + fp) == 0) NA_real_ else tp / (tp + fp)
  recall    <- if ((tp + fn) == 0) NA_real_ else tp / (tp + fn)
  f1        <- if (is.na(precision) || is.na(recall) || (precision + recall) == 0) {
    NA_real_
  } else {
    2 * precision * recall / (precision + recall)
  }
  data.frame(
    Class     = cls,
    Precision = precision,
    Recall    = recall,
    F1        = f1
  )
})
metrics_df <- bind_rows(metrics_list)

metrics_long <- metrics_df %>%
  pivot_longer(cols = c("Precision", "Recall", "F1"),
               names_to = "Metric",
               values_to = "Value")

ggplot(metrics_long, aes(x = Class, y = Value, fill = Metric)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 1)
  ) +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Per-Class Precision, Recall, and F1 (Test Set)",
    x = "Sentiment Class",
    y = "Score"
  ) +
  theme_minimal()

```

#4.8 Distrubtion of true & predicted sentiment of the test set (in bar chart)

```{r}

true_df <- data.frame(Label = test_labels, Type = "True")

pred_df <- data.frame(Label = pred_test,   Type = "Predicted")

dist_df <- rbind(true_df, pred_df)

ggplot(dist_df, aes(x = Label, fill = Type)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = c("True" = "#3182bd", "Predicted" = "#e6550d")) +
  labs(
    title = "Distribution of True vs Predicted Sentiment (Test Set)",
    x = "Sentiment Class",
    y = "Count",
    fill = NULL
  ) +
  theme_minimal()

```
